---
title: "Data Cleaning"
output: html_document
date: 2025-02-26
---

# Load packages
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(janitor)
library(tidycensus)

 numberT)
```

# Template
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))
 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#Load and Clean EQI Data
```{r}
#Load EQI data
eqi_raw <- read_csv("data/2006_2010_EQI_2Jan2018_VC.csv") |> 
  clean_names(case = "snake")

head(eqi_raw)
colnames(eqi_raw)

#Add leading zeros to the stFIPS; code taken from the provided EQI README file 
eqi_raw$stfips <- substr(as.numeric(eqi_raw$stfips) + 100000, 2,7)
eqi_raw$stfips

#Rename variables
eqi_df =
  eqi_raw |>
  rename(
    fips = stfips,
    state_abbr = state,
    air = air_eqi_2jan2018_vc,
    air_rucc1 = rucc1_air_eqi_2jan2018_vc,
    air_rucc2 = rucc2_air_eqi_2jan2018_vc,
    air_rucc3 = rucc3_air_eqi_2jan2018_vc,
    air_rucc4 = rucc4_air_eqi_2jan2018_vc,
    air_4q = air_eqi_2jan2018_vc_4,
    air_4q_rucc1 = rucc1_air_eqi_2jan2018_vc_4,
    air_4q_rucc2 = rucc2_air_eqi_2jan2018_vc_4,
    air_4q_rucc3 = rucc3_air_eqi_2jan2018_vc_4,
    air_4q_rucc4 = rucc4_air_eqi_2jan2018_vc_4,
    air_3q = air_eqi_2jan2018_vc_3,
    air_3q_rucc1 = rucc1_air_eqi_2jan2018_vc_3,
    air_3q_rucc2 = rucc2_air_eqi_2jan2018_vc_3,
    air_3q_rucc3 = rucc3_air_eqi_2jan2018_vc_3,
    air_3q_rucc4 = rucc4_air_eqi_2jan2018_vc_3,
    air_5q = air_eqi_2jan2018_vc_5,
    air_5q_rucc1 = rucc1_air_eqi_2jan2018_vc_5,
    air_5q_rucc2 = rucc2_air_eqi_2jan2018_vc_5,
    air_5q_rucc3 = rucc3_air_eqi_2jan2018_vc_5,
    air_5q_rucc4 = rucc4_air_eqi_2jan2018_vc_5,
    water = water_eqi_2jan2018_vc,
    water_rucc1 = rucc1_water_eqi_2jan2018_vc,
    water_rucc2 = rucc2_water_eqi_2jan2018_vc,
    water_rucc3 = rucc3_water_eqi_2jan2018_vc,
    water_rucc4 = rucc4_water_eqi_2jan2018_vc,
    water_4q = water_eqi_2jan2018_vc_4,
    water_4q_rucc1 = rucc1_water_eqi_2jan2018_vc_4,
    water_4q_rucc2 = rucc2_water_eqi_2jan2018_vc_4,
    water_4q_rucc3 = rucc3_water_eqi_2jan2018_vc_4,
    water_4q_rucc4 = rucc4_water_eqi_2jan2018_vc_4,
    water_3q = water_eqi_2jan2018_vc_3,
    water_3q_rucc1 = rucc1_water_eqi_2jan2018_vc_3,
    water_3q_rucc2 = rucc2_water_eqi_2jan2018_vc_3,
    water_3q_rucc3 = rucc3_water_eqi_2jan2018_vc_3,
    water_3q_rucc4 = rucc4_water_eqi_2jan2018_vc_3,
    water_5q = water_eqi_2jan2018_vc_5,
    water_5q_rucc1 = rucc1_water_eqi_2jan2018_vc_5,
    water_5q_rucc2 = rucc2_water_eqi_2jan2018_vc_5,
    water_5q_rucc3 = rucc3_water_eqi_2jan2018_vc_5,
    water_5q_rucc4 = rucc4_water_eqi_2jan2018_vc_5,
    land = land_eqi_2jan2018_vc,
    land_rucc1 = rucc1_land_eqi_2jan2018_vc,
    land_rucc2 = rucc2_land_eqi_2jan2018_vc,
    land_rucc3 = rucc3_land_eqi_2jan2018_vc,
    land_rucc4 = rucc4_land_eqi_2jan2018_vc,
    land_4q = land_eqi_2jan2018_vc_4,
    land_4q_rucc1 = rucc1_land_eqi_2jan2018_vc_4,
    land_4q_rucc2 = rucc2_land_eqi_2jan2018_vc_4,
    land_4q_rucc3 = rucc3_land_eqi_2jan2018_vc_4,
    land_4q_rucc4 = rucc4_land_eqi_2jan2018_vc_4,
    land_3q = land_eqi_2jan2018_vc_3,
    land_3q_rucc1 = rucc1_land_eqi_2jan2018_vc_3,
    land_3q_rucc2 = rucc2_land_eqi_2jan2018_vc_3,
    land_3q_rucc3 = rucc3_land_eqi_2jan2018_vc_3,
    land_3q_rucc4 = rucc4_land_eqi_2jan2018_vc_3,
    land_5q = land_eqi_2jan2018_vc_5,
    land_5q_rucc1 = rucc1_land_eqi_2jan2018_vc_5,
    land_5q_rucc2 = rucc2_land_eqi_2jan2018_vc_5,
    land_5q_rucc3 = rucc3_land_eqi_2jan2018_vc_5,
    land_5q_rucc4 = rucc4_land_eqi_2jan2018_vc_5,
    sociod = sociod_eqi_2jan2018_vc,
    sociod_rucc1 = rucc1_sociod_eqi_2jan2018_vc,
    sociod_rucc2 = rucc2_sociod_eqi_2jan2018_vc,
    sociod_rucc3 = rucc3_sociod_eqi_2jan2018_vc,
    sociod_rucc4 = rucc4_sociod_eqi_2jan2018_vc,
    sociod_4q = sociod_eqi_2jan2018_vc_4,
    sociod_4q_rucc1 = rucc1_sociod_eqi_2jan2018_vc_4,
    sociod_4q_rucc2 = rucc2_sociod_eqi_2jan2018_vc_4,
    sociod_4q_rucc3 = rucc3_sociod_eqi_2jan2018_vc_4,
    sociod_4q_rucc4 = rucc4_sociod_eqi_2jan2018_vc_4,
    sociod_3q = sociod_eqi_2jan2018_vc_3,
    sociod_3q_rucc1 = rucc1_sociod_eqi_2jan2018_vc_3,
    sociod_3q_rucc2 = rucc2_sociod_eqi_2jan2018_vc_3,
    sociod_3q_rucc3 = rucc3_sociod_eqi_2jan2018_vc_3,
    sociod_3q_rucc4 = rucc4_sociod_eqi_2jan2018_vc_3,
    sociod_5q = sociod_eqi_2jan2018_vc_5,
    sociod_5q_rucc1 = rucc1_sociod_eqi_2jan2018_vc_5,
    sociod_5q_rucc2 = rucc2_sociod_eqi_2jan2018_vc_5,
    sociod_5q_rucc3 = rucc3_sociod_eqi_2jan2018_vc_5,
    sociod_5q_rucc4 = rucc4_sociod_eqi_2jan2018_vc_5,
    built = built_eqi_2jan2018_vc,
    built_rucc1 = rucc1_built_eqi_2jan2018_vc,
    built_rucc2 = rucc2_built_eqi_2jan2018_vc,
    built_rucc3 = rucc3_built_eqi_2jan2018_vc,
    built_rucc4 = rucc4_built_eqi_2jan2018_vc,
    built_4q = built_eqi_2jan2018_vc_4,
    built_4q_rucc1 = rucc1_built_eqi_2jan2018_vc_4,
    built_4q_rucc2 = rucc2_built_eqi_2jan2018_vc_4,
    built_4q_rucc3 = rucc3_built_eqi_2jan2018_vc_4,
    built_4q_rucc4 = rucc4_built_eqi_2jan2018_vc_4,
    built_3q = built_eqi_2jan2018_vc_3,
    built_3q_rucc1 = rucc1_built_eqi_2jan2018_vc_3,
    built_3q_rucc2 = rucc2_built_eqi_2jan2018_vc_3,
    built_3q_rucc3 = rucc3_built_eqi_2jan2018_vc_3,
    built_3q_rucc4 = rucc4_built_eqi_2jan2018_vc_3,
    built_5q = built_eqi_2jan2018_vc_5,
    built_5q_rucc1 = rucc1_built_eqi_2jan2018_vc_5,
    built_5q_rucc2 = rucc2_built_eqi_2jan2018_vc_5,
    built_5q_rucc3 = rucc3_built_eqi_2jan2018_vc_5,
    built_5q_rucc4 = rucc4_built_eqi_2jan2018_vc_5,
    eqi = eqi_2jan2018_vc,
    eqi_rucc1 = rucc1_eqi_2jan2018_vc,
    eqi_rucc2 = rucc2_eqi_2jan2018_vc,
    eqi_rucc3 = rucc3_eqi_2jan2018_vc,
    eqi_rucc4 = rucc4_eqi_2jan2018_vc,
    eqi_4q = eqi_2jan2018_vc_4,
    eqi_4q_rucc1 = rucc1_eqi_2jan2018_vc_4,
    eqi_4q_rucc2 = rucc2_eqi_2jan2018_vc_4,
    eqi_4q_rucc3 = rucc3_eqi_2jan2018_vc_4,
    eqi_4q_rucc4 = rucc4_eqi_2jan2018_vc_4,
    eqi_3q = eqi_2jan2018_vc_3,
    eqi_3q_rucc1 = rucc1_eqi_2jan2018_vc_3,
    eqi_3q_rucc2 = rucc2_eqi_2jan2018_vc_3,
    eqi_3q_rucc3 = rucc3_eqi_2jan2018_vc_3,
    eqi_3q_rucc4 = rucc4_eqi_2jan2018_vc_3,
    eqi_5q = eqi_2jan2018_vc_5,
    eqi_5q_rucc1 = rucc1_eqi_2jan2018_vc_5,
    eqi_5q_rucc2 = rucc2_eqi_2jan2018_vc_5,
    eqi_5q_rucc3 = rucc3_eqi_2jan2018_vc_5,
    eqi_5q_rucc4 = rucc4_eqi_2jan2018_vc_5
  )

dim(eqi_df)
#3143 counties are captured in the eqi_df

sum(duplicated(eqi_df$fips))
#no duplicate counties in the eqi_df

length(unique(eqi_df$state_abbr))
unique(eqi_df$state_abbr)
#51 states because it includes DC

sum(is.na(eqi_df$eqi))
sum(is.na(eqi_df$air))
sum(is.na(eqi_df$water))
sum(is.na(eqi_df$land))
sum(is.na(eqi_df$sociod))
#no NA values for our main predictors
```

#Load and Clean County Health Data
```{r}
#Load county health data; clean up names and select interested variables
ch_raw <- read_xlsx(path = "data/2015 County Health Rankings Data.xlsx", sheet = 4, 
                   skip =1, range = "A2:FE3143")

colnames(ch_raw)

ch_df <- ch_raw |> 
  rename(fips = "FIPS",
         state_name = "State",
         county_abbr = "County",
         unrealiable_bw = "Unreliable",
         num_low_birthweight_births = "# Low Birthweight Births",
         num_live_births = "# Live births",
         perc_low_bw = "% LBW",
         perc_smokers = "% Smokers") |> 
  select(fips, state_name, county_abbr, unrealiable_bw, num_low_birthweight_births, 
         num_live_births, perc_low_bw, perc_smokers)

ch_raw2 =
  read_xlsx("data/2015 County Health Rankings Data.xlsx", sheet = 6, skip = 1, range = "A2:CB3143")

ch_df2 <- ch_raw2 |> 
  rename(fips = "FIPS",
         state_name = "State",
         county_abbr = "County",
         perc_black = "% African American",
         perc_AIAN = "% American Indian/ Alaskan Native",
         perc_asian = "% Asian",
         perc_islander = "% Native Hawaiian/ Other Pacific Islander",
         perc_hisp = "% Hispanic",
         perc_white = "% Non-Hispanic white",
         perc_rural = "% Rural") |> 
  select(fips, state_name, county_abbr, perc_black, perc_AIAN, perc_asian, perc_islander, perc_hisp, perc_white, perc_rural)

ch_df =
  left_join(ch_df, ch_df2)

dim(ch_df)
#county health data captures only 3141 counties

sum(duplicated(ch_df$fips))
#no duplicate counties

length(unique(ch_df$state_name))
unique(ch_df$state_name)
#51 states because it includes DC
```

# Loading and cleaning census data for age
```{r}
# importing variable list and selecting variables of interest from census

variables = load_variables(2013, "acs5")

age_data = get_acs(
  year = 2013,
  geography = "county",
  variables = c("B01001_026", "B01001_031", "B01001_032",
                "B01001_033", "B01001_034", "B01001_035",
                "B01001_036", "B01001_037", "B01001_038",
                "B01001_039"),
         geometry = TRUE,
         survey = "acs5",
         output = "wide") |>
  rename(total_age = B01001_026E,
         t1819 = B01001_031E,
         t20 = B01001_032E,
         t21 = B01001_033E,
         t2224 = B01001_034E,
         t2529 = B01001_035E,
         t3034 = B01001_036E,
         t3539 = B01001_037E,
         t4044 = B01001_038E,
         t4549 = B01001_039E) |> 
  mutate(age_18_49 = t1819+t20+t21+t2224+t2529+t3034+t3539+t4044+t4549,
         age_18_24 = t1819+t20+t21+t2224,
         age_25_34 = t2529+t3034,
         age_35_44 = t3539+t4044,
         age_45_49 = t4549) |> 
  mutate(perc_age1849 = 100*(age_18_49/total_age),
         perc_age1824 = 100*(age_18_24/total_age),
         perc_age2534 = 100*(age_25_34/total_age),
         perc_age3544 = 100*(age_35_44/total_age),
         perc_age4549 = 100*(age_45_49/total_age)) |> 
  janitor::clean_names() |> 
  select(name, geoid, perc_age1849, perc_age1824, perc_age2534, perc_age3544, perc_age4549) |> 
  rename(fips = geoid)

ch_df =
  left_join(ch_df, age_data) |> 
  select(-name)
```


#Create combined dataset
```{r}
#combine datasets
eqi_lbw_raw_df <- inner_join(eqi_df, ch_df, by = "fips")

#create offset of log_live_births
#reorder columns
eqi_lbw_raw_df <- eqi_lbw_raw_df |> 
  mutate(log_live_births = log(num_live_births)) |> 
  select(fips, state_name, state_abbr, county_name, county_abbr, 
         unrealiable_bw, num_low_birthweight_births, num_live_births, 
         log_live_births, perc_low_bw, everything())

dim(eqi_lbw_raw_df)
#when joining datasets together get only 3138 counties

#which counties are missing?
#anti_join returns all rows from x without a match in y
anti <- anti_join(eqi_df, ch_df, by = "fips")
anti <- anti |> select(fips, everything())
#5 counties in Alaska are in eqi_df but not ch_df

#anti_join returns all rows from x without a match in y
anti2 <- anti_join(ch_df, eqi_df, by = "fips")
anti2 <- anti2 |> select(fips, everything())
#3 counties in Alaska are in ch_df but not eqi_df

```
We have 3138 counties captured in both datasets.

#Check for NA and unreliable lbw values
#unreliable means value reported but considered unreliable since based on counts of twenty or less
```{r}
sum(is.na(eqi_lbw_raw_df$num_low_birthweight_births))
#99 counties are missing num_low_birthweight_births

#which states have most missing num_low_birthweight_births?
eqi_lbw_raw_df |> 
  filter(is.na(num_low_birthweight_births)) |> 
  group_by(state_name) |> 
  summarize(num_county_na = n()) |> 
  arrange(desc(num_county_na))
#Nebraska has the most missing low bw values

sum(is.na(eqi_lbw_raw_df$num_live_births))
#99 counties are missing num_live_births

table(eqi_lbw_raw_df$unrealiable_bw)
#the x marks the counties with unreliable birth weight data
#44 counties have unreliable low birthweight data

#which states have most number of unreliable num_low_birthweight_births?
eqi_lbw_raw_df |> 
  filter(unrealiable_bw %in% c("x")) |> 
  group_by(state_name) |> 
  summarize(num_county_na = n()) |> 
  arrange(desc(num_county_na))
#Kansas has the most missing unreliable values

```

#Create cleaned dataset with unreliable and and NA values removed for low bw
```{r}
eqi_lbw_clean_df <- eqi_lbw_raw_df |> filter(! unrealiable_bw %in% c("x"))

eqi_lbw_clean_df <- eqi_lbw_clean_df |> 
  filter(! is.na(num_low_birthweight_births)) |> 
  select(-unrealiable_bw)

dim(eqi_lbw_clean_df)
#Our cleaned dataset has 2995 counties

sum(is.na(eqi_lbw_clean_df$num_low_birthweight_births))
sum(is.na(eqi_lbw_clean_df$num_live_births))
sum(is.na(eqi_lbw_clean_df$log_live_births))
#checked and have no NA values for these variables

```
Our cleaned dataset has 2995 counties

#Write the combined dataset
```{r}
#write_csv(eqi_lbw_raw_df,"data/eqi_lbw_raw_df.csv")

#write_csv(eqi_lbw_clean_df,"data/eqi_lbw_clean_df.csv")

```
raw dataset has unreliable low bw values and the NA low bw values while cleaned dataset does not
