---
title: "EDA"
output: html_document
date: 2025-02-26
---

# Load packages
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(splines)
library(mgcv)
library(ggcorrplot)
library(table1)
#library(xlsx)
library(cowplot)
library(ggplot2)
```

# Template
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))
 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

# Load dataset
Clean dataset has unreliable low bw values and the NA low bw values removed
```{r}
eqi_lbw_clean_df <- read_csv("data/eqi_lbw_clean_df.csv")

```

# Explore counties
```{r}
dim(eqi_lbw_clean_df)
#we have 2995 counties in total

#how many counties do we have for each state?
eqi_lbw_clean_df |> 
  group_by(state_name) |> 
  summarize(num_county = n()) |> 
  ggplot(aes(x = state_name, y = num_county)) +
  geom_col()
#NEED TO CLEAN UP THIS GRAPH

#which state has the most counties?
eqi_lbw_clean_df |> 
  group_by(state_name) |> 
  summarize(num_county = n()) |> 
  arrange(desc(num_county))
#Texas has the most counties described here; 235 counties

#which state has the least counties?
eqi_lbw_clean_df |> 
  group_by(state_name) |> 
  summarize(num_county = n()) |> 
  arrange(num_county)
#DC has the least counties described here; 1 county
#next is DE with 3 counties described
#there are no states with 0 counties described

```

# Distributions
## num_low_birthweight_births
```{r}
sum(is.na(eqi_lbw_clean_df$num_low_birthweight_births))
#removed NA values previously

eqi_lbw_clean_df |> ggplot(aes(x = num_low_birthweight_births)) + 
  geom_histogram() 

#zoom in
eqi_lbw_clean_df |> ggplot(aes(x = num_low_birthweight_births)) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 2000))  # Set x-axis range

eqi_lbw_clean_df |> 
  select(county_name, state_name, num_low_birthweight_births) |> 
  arrange(desc(num_low_birthweight_births)) |> 
  print(n=50)
#Los Angeles County, California has highest num_low_birthweight_births = 71362

eqi_lbw_clean_df |> 
  select(county_name, state_name, num_low_birthweight_births) |> 
  arrange(num_low_birthweight_births) |> 
  print(n=50)
#Sedgwick County, Colorado has lowest num_low_birthweight_births = 10

#range is 10 to 71362 counts of low birthweight

mean(eqi_lbw_clean_df$num_low_birthweight_births)
median(eqi_lbw_clean_df$num_low_birthweight_births)
IQR(eqi_lbw_clean_df$num_low_birthweight_births)
#num_low_birthweight_births is incredibly right skewed

```
num_low_birthweight_births is incredibly right skewed

## num_live_births
```{r}
sum(is.na(eqi_lbw_clean_df$num_live_births))
#removed NA values previously

eqi_lbw_clean_df |> ggplot(aes(x = num_live_births)) + 
  geom_histogram() 

#zoom in
eqi_lbw_clean_df |> ggplot(aes(x = num_live_births)) + 
  geom_histogram() + 
  scale_x_continuous(limits = c(0, 2000))  # Set x-axis range

eqi_lbw_clean_df |> 
  select(county_name, state_name, num_live_births) |> 
  arrange(desc(num_live_births)) |> 
  print(n=50)
#Los Angeles County, California has highest num_live_births = 986350

eqi_lbw_clean_df |> 
  select(county_name, state_name, num_live_births) |> 
  arrange(num_live_births) |> 
  print(n=50)
#Jackson County, Colorado has lowest num_live_births = 77

#range is 77 to 986350 counts for num_live_births

mean(eqi_lbw_clean_df$num_live_births, na.rm = TRUE)
median(eqi_lbw_clean_df$num_live_births, na.rm = TRUE)
IQR(eqi_lbw_clean_df$num_live_births, na.rm = TRUE)
#live birth counts are incredibly right skewed

```
live birth counts are incredibly right skewed

## EQI - higher means poorer environment
```{r}
sum(is.na(eqi_lbw_clean_df$eqi))
#no NAs

eqi_lbw_clean_df |> ggplot(aes(x = eqi)) + geom_histogram()

eqi_lbw_clean_df |> 
  select(county_name, state_name, eqi) |> 
  arrange(desc(eqi))

eqi_lbw_clean_df |> 
  select(county_name, state_name, eqi) |> 
  arrange(eqi)

```
Eqi is mostly, normal. Slight left tail.

## Sociod
```{r}
sum(is.na(eqi_lbw_clean_df$sociod))
#no NAs

eqi_lbw_clean_df |> ggplot(aes(x = sociod)) + geom_histogram()

eqi_lbw_clean_df |> 
  select(county_name, state_name, sociod) |> 
  arrange(desc(sociod))

eqi_lbw_clean_df |> 
  select(county_name, state_name, sociod) |> 
  arrange(sociod)

```
Sociod is Left skewed.

## Air
```{r}
sum(is.na(eqi_lbw_clean_df$air))
#no NAs

eqi_lbw_clean_df |> ggplot(aes(x = air)) + geom_histogram()

eqi_lbw_clean_df |> 
  select(county_name, state_name, air) |> 
  arrange(desc(sociod))

eqi_lbw_clean_df |> 
  select(county_name, state_name, air) |> 
  arrange(sociod)

```
Air is normally distributed.

## Water - distribution is bad
```{r}
sum(is.na(eqi_lbw_clean_df$water))
#no NAs

eqi_lbw_clean_df |> ggplot(aes(x = water)) + geom_histogram()

eqi_lbw_clean_df |> 
  select(county_name, state_name, water) |> 
  arrange(desc(water))

eqi_lbw_clean_df |> 
  select(county_name, state_name, water) |> 
  arrange(water)

```
Water is Bad. Use a cutoff?

## Land
```{r}
sum(is.na(eqi_lbw_clean_df$land))
#no NAs

eqi_lbw_clean_df |> ggplot(aes(x = land)) + geom_histogram()

eqi_lbw_clean_df |> 
  select(county_name, state_name, land) |> 
  arrange(desc(land))

eqi_lbw_clean_df |> 
  select(county_name, state_name, land) |> 
  arrange(land)

```
Land is left skewed

## Built
```{r}
sum(is.na(eqi_lbw_clean_df$built))
#no NAs

eqi_lbw_clean_df |> ggplot(aes(x = built)) + geom_histogram()

eqi_lbw_clean_df |> 
  select(county_name, state_name, built) |> 
  arrange(desc(built))

eqi_lbw_clean_df |> 
  select(county_name, state_name, built) |> 
  arrange(built)

```
Built is normally distributed.

## num_low_birthweight_births and EQI

```{r}
eqi_lbw_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = eqi, y = num_low_birthweight_births)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Scatterplot of EQI and Low Birth Weight Births",
        x = "Environmental Quality Index (EQI)",
        y = "Number of Low Birth Weight Births") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

# ggsave("EDA_files/EQI_LBW_Distribution.png", eqi_lbw_plot, width = 8, height = 4, dpi = 300)
```

```{r}
lbw_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = num_low_birthweight_births)) +
  geom_histogram() +
  labs(title = "Distribution of Low Birth Weight Births",
        x = "Number of Low Birth Weight Births",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

eqi_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = eqi)) +
  geom_histogram() +
  labs(title = "Distribution of Environmental Quality Index (EQI)",
        x = "Environmental Quality Index (EQI)",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

eqi_lbw_hist_plot = plot_grid(lbw_plot, eqi_plot, ncol = 2)

# ggsave("EDA_files/EQI_LBW_Histograms.png", eqi_lbw_hist_plot, width = 10, height = 4, dpi = 300)
```

```{r}
air_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = air)) +
  geom_histogram() +
  labs(title = "Distribution of Air Index",
        x = "Air Index",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

land_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = land)) +
  geom_histogram() +
  labs(title = "Distribution of Land Index",
        x = "Land Index",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

water_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = water)) +
  geom_histogram() +
  labs(title = "Distribution of Water Index",
        x = "Water Index",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

built_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = built)) +
  geom_histogram() +
  labs(title = "Distribution of Built Index",
        x = "Built Index",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 

sociod_plot =
  eqi_lbw_clean_df |> 
  ggplot(aes(x = air)) +
  geom_histogram() +
  labs(title = "Distribution of Sociodemographic Index",
        x = "Sociodemographic Index",
        y = "Count") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) 


sub_eqi_plot = plot_grid(air_plot, water_plot, land_plot, built_plot, sociod_plot, ncol = 2)

# ggsave("EDA_files/Sub_EQI_Histograms.png", sub_eqi_plot, width = 10, height = 6, dpi = 300)
```

# Quick Look at Correlations (using Spearman)
```{r}
cor_var <- eqi_lbw_clean_df |> 
  select(eqi, water, 
         built, air,
         land, sociod,
         num_low_birthweight_births,
         num_live_births,
         perc_smokers, perc_rural,
         perc_age1849, perc_age1824, perc_age2534, perc_age3544, perc_age4549) |> 
  na.omit()

##Spearman correlations
corr <- round(cor(cor_var, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(cor_var)

#simple plot
ggcorrplot(corr)

#hiearchial clustering
ggcorrplot(corr, hc.order = TRUE, outline.color = "white")

#lower triangle
ggcorrplot(corr,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white")

#add correlation values onto plot
ggcorrplot(corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)

```

# Confounder Exploration - examining rural, smoking, reproductive age for females
```{r}
###perc smokers across counties
eqi_lbw_clean_df |> ggplot(aes(x = perc_smokers)) + geom_histogram()
#slight right tail

#which counties have the highest perc_smokers?
eqi_lbw_clean_df |> 
  arrange(desc(perc_smokers))

#which counties have the lowest perc_smokers?
eqi_lbw_clean_df |> 
  arrange(perc_smokers)

###perc rural
eqi_lbw_clean_df |> ggplot(aes(x = perc_rural)) + geom_histogram()
#super left skewed; pretty similar than large spike at 100%

#which counties have the highest perc_rural?
rural_counties <- eqi_lbw_clean_df |> 
  arrange(desc(perc_rural)) |> 
  select(state_name, county_name, perc_rural)

#view(rural_counties)

#which counties have the lowest perc_rural?
eqi_lbw_clean_df |> 
  arrange(perc_smokers)

###rural as a categorical variable from eqi dataset
#make rucc categories a factor
eqi_lbw_clean_df <- eqi_lbw_clean_df |> 
  mutate(cat_rucc = factor(cat_rucc, levels = c("1","2","3", "4"),labels = c("metropolitan-urbanized",
                                                                      "nonmetropolitan-urbanized",
                                                                      "less urbanized",
                                                                      "thinly populated")))

num_counties <- nrow(eqi_lbw_clean_df)

rucc_cat_summary <- eqi_lbw_clean_df |> 
  count(cat_rucc) |> 
  mutate(prop = n/num_counties*100)

eqi_lbw_clean_df |> 
  ggplot(aes(x = cat_rucc)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

###perc_age1849 across counties (entire reproductive age range for females)
eqi_lbw_clean_df |> ggplot(aes(x = perc_age1849)) + geom_histogram()
#slight right tail

###perc_age1824 for females across counties
eqi_lbw_clean_df |> ggplot(aes(x = perc_age1824)) + geom_histogram()
#right tail

###perc_age2534 for females across counties
eqi_lbw_clean_df |> ggplot(aes(x = perc_age2534)) + geom_histogram()
#slight right tail

###perc_age3544 for females across counties
eqi_lbw_clean_df |> ggplot(aes(x = perc_age3544)) + geom_histogram()
#normally distributed

###perc_age4549 females across counties
eqi_lbw_clean_df |> ggplot(aes(x = perc_age4549)) + geom_histogram()
#slight right tail

```

#Simple quasipoisson regression of Low birthweight counts (outcome) with each individual covariate
```{r}
simple_reg_lbw <- 
  expand_grid(
    outcome_name = c("num_low_birthweight_births"),
    var_names = c("eqi", "perc_smokers", "perc_rural", "perc_age1849", "perc_age1824", 
                  "perc_age2534","perc_age3544", "perc_age4549", "land", "air", "sociod", 
                  "built", "water", "cat_rucc")
  ) |> 
  mutate(models = map2(outcome_name, var_names, \(o,v) glm(paste0(o, "~", v), data = eqi_lbw_clean_df, 
                                                           family = quasipoisson, 
                                                           offset = log_live_births)),
         result = map(models, broom::tidy, conf.int=TRUE)) |> 
  unnest(result) |> 
  filter(term != "(Intercept)")

simple_reg_signif_lbw <- simple_reg_lbw |> 
  select(outcome_name, var_names, term, estimate, p.value, conf.low, conf.high) |> 
  filter(p.value <= 0.05)

view(simple_reg_signif_lbw)

```

#Simple linear regression of EQI (exposure) with each individual covariate
```{r}
simple_reg_eqi <- 
  expand_grid(
    exp_name = c("eqi"),
    var_names = c("perc_smokers", "perc_rural", "perc_age1849", "perc_age1824", 
                  "perc_age2534","perc_age3544", "perc_age4549", "cat_rucc")
  ) |> 
  mutate(models = map2(exp_name, var_names, \(e,v) lm(paste0(e, "~", v), data = eqi_lbw_clean_df)),
         result = map(models, broom::tidy, conf.int=TRUE)) |> 
  unnest(result) |> 
  filter(term != "(Intercept)")

simple_reg_signif_eqi <- simple_reg_eqi |> 
  select(exp_name, var_names, term, estimate, p.value, conf.low, conf.high) |> 
  filter(p.value <= 0.05)

view(simple_reg_signif_eqi)

```

#Simple linear regression of each subdomain (secondary exposure) with each individual covariate
##CLEAN THIS up by domain
```{r}
simple_reg_domains <- 
  expand_grid(
    exp_name = c("air", "land", "water", "built", "sociod"),
    var_names = c("perc_smokers", "perc_rural", "perc_age1849", "perc_age1824", 
                  "perc_age2534","perc_age3544", "perc_age4549", "air", "land", 
                  "water", "built", "sociod")
  ) |> 
  mutate(models = map2(exp_name, var_names, \(e,v) lm(paste0(e, "~", v), data = eqi_lbw_clean_df)),
         result = map(models, broom::tidy, conf.int=TRUE)) |> 
  unnest(result) |> 
  filter(term != "(Intercept)")

simple_reg_signif_domains <- simple_reg_domains |> 
  select(exp_name, var_names, term, estimate, p.value, conf.low, conf.high) |> 
  filter(p.value <= 0.05)

view(simple_reg_signif_domains)

```

# Sumamry Tables

```{r, message = F}
# creating a separate dataset for the summary tables with clearer names and formatting

tbl_data =
  eqi_lbw_clean_df |> 
  mutate(`Environmental Quality Index` = eqi,
          eqi_cat =  cut(eqi, 
                  breaks = quantile(eqi, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
                  labels = c("Excellent", "Good", "Fair", "Poor"),
                  include.lowest = TRUE),
         `Air Domain Index` = air,
          air_cat =  cut(air, 
                  breaks = quantile(air, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
                  labels = c("Excellent", "Good", "Fair", "Poor"),
                  include.lowest = TRUE),
         `Water Domain Index` = water,
          water_cat =  cut(water, 
                  breaks = quantile(water, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
                  labels = c("Excellent", "Good", "Fair", "Poor"),
                  include.lowest = TRUE),
         `Land Domain Index` = land,
          land_cat =  cut(land, 
                  breaks = quantile(land, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
                  labels = c("Excellent", "Good", "Fair", "Poor"),
                  include.lowest = TRUE),
         `Built-Environment Domain Index` = built,
          built_cat =  cut(built, 
                  breaks = quantile(built, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
                  labels = c("Excellent", "Good", "Fair", "Poor"),
                  include.lowest = TRUE),
         `Sociodemographic Domain Index` = sociod,
          sociod_cat =  cut(sociod, 
                  breaks = quantile(sociod, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE),
                  labels = c("Excellent", "Good", "Fair", "Poor"),
                  include.lowest = TRUE),
         `Rural-Urban Continuum Codes` = case_when(cat_rucc == 1 ~ "Metropolitan-Urbanized",
                                                   cat_rucc == 2 ~ "Nonmetropolitan-Urbanized ",
                                                   cat_rucc == 3 ~ "Less Urbanized",
                                                   cat_rucc == 4 ~ "Thinly Populated"),
         `Live Births (#)` = num_live_births,
         `Low Birth Weight (#)` = num_low_birthweight_births,
         `Low Birth Weight (%)` = perc_low_bw,
         `Non-Hispanic White (%)` = perc_white,
         `African-American (%)` = perc_black,
         `Asian (%)` = perc_asian,
         `American Indian / Alaskan Native (%)` = perc_AIAN,
         `Native Hawaiian / Other Pacific Islander (%)` = perc_islander,
         `Hispanic (%)` = perc_hisp,
         `18-49 years old (%)` = perc_age1849, 
         `18-24 years old (%)` = perc_age1824, 
         `25-34 years old (%)` = perc_age2534, 
         `35-44 years old (%)` = perc_age3544, 
         `45-49 years old (%)` = perc_age4549,
         `Rural (%)` = perc_rural,
         `Smokers (%)` = perc_smokers)
```

## Table 1

```{r, message = F, warning = F}
# creating a table 1 - overall

table1 =
  table1::table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)`, data = tbl_data)

table1

table1_df = data.frame(table1)
```

## Table 2

```{r, message = F, warning = F}
# creating a table 2 based on the overall EQI

table2_eqi = 
  table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)` | eqi_cat, data = tbl_data, caption = "Table 1: Demographic Characteristics by Environmental Quality Index Category per County in 2015")

table2_eqi

table2_eqi = data.frame(table2_eqi)
```


```{r, message = F, warning = F}
# creating a table 2 based on the air index

table2_air = 
  table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)` | air_cat, data = tbl_data)

table2_air

table2_air = data.frame(table2_air)
```

```{r, message = F, warning = F}
# creating a table 2 based on the water index

table2_water = 
  table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)` | water_cat, data = tbl_data)

table2_water

table2_water = data.frame(table2_water)
```

```{r, message = F, warning = F}
# creating a table 2 based on the land index

table2_land = 
  table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)` | land_cat, data = tbl_data)

table2_land

table2_land = data.frame(table2_land)
```

```{r, message = F, warning = F}
# creating a table 2 based on the built index

table2_built = 
  table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)` | built_cat, data = tbl_data)

table2_built

table2_built = data.frame(table2_built)
```


```{r, message = F, warning = F}
# creating a table 2 based in the sociodemographic index

table2_sociod = 
  table1(~ `Environmental Quality Index` + `Air Domain Index` + `Water Domain Index` + `Land Domain Index` + `Built-Environment Domain Index` + `Sociodemographic Domain Index` + `Rural-Urban Continuum Codes` + `Live Births (#)` + `Low Birth Weight (#)` + `Low Birth Weight (%)` + `Non-Hispanic White (%)` + `African-American (%)` + `Asian (%)` + `American Indian / Alaskan Native (%)` + `Native Hawaiian / Other Pacific Islander (%)` + `Hispanic (%)` + `18-49 years old (%)` + `18-24 years old (%)` + `25-34 years old (%)` + `35-44 years old (%)` + `45-49 years old (%)` + `Rural (%)` + `Smokers (%)` | sociod_cat, data = tbl_data)

table2_sociod

table2_sociod = data.frame(table2_sociod)
```

```{r, eval = F}
# exporting the tables into one excel sheet

wb = createWorkbook()

sheet = createSheet(wb, "overall")
xlsx::addDataFrame(table1, sheet=sheet, startColumn=1, row.names=FALSE)

sheet = createSheet(wb, "eqi")
xlsx::addDataFrame(table2_eqi, sheet=sheet, startColumn=1, row.names=FALSE)

sheet = xlsx::createSheet(wb, "air")
xlsx::addDataFrame(table2_air, sheet=sheet, startColumn=1, row.names=FALSE)

sheet = xlsx::createSheet (wb, "water")
xlsx::addDataFrame(table2_water, sheet=sheet, startColumn=1, row.names=FALSE)

sheet = xlsx::createSheet (wb, "land")
xlsx:: addDataFrame(table2_land, sheet=sheet, startColumn=1, row.names=FALSE)

sheet = xlsx::createSheet (wb, "built")
xlsx:: addDataFrame(table2_built, sheet=sheet, startColumn=1, row.names=FALSE)

sheet = xlsx::createSheet (wb, "sociod")
xlsx:: addDataFrame(table2_sociod, sheet=sheet, startColumn=1, row.names=FALSE)

xlsx::saveWorkbook (wb=wb,"data/summary_tables.xlsx")
```