---
title: "Prelim Models"
output: html_document
date: 2025-02-26
---

# Load packages
```{r}
library(tidyverse)
library(readr)
library(readxl)
library(splines)
library(mgcv)

```

# Template
```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))
 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

#Load dataset
Clean dataset has unreliable low bw values and the NA low bw values removed
```{r}
eqi_lbw_clean_df <- read_csv("data/eqi_lbw_clean_df.csv")

```

#Prelim model just with EQI as exposure and low bw as outcome
##Poisson model
```{r}
######Fit a Poisson model (not accounting for overdispersion)
mod1_p = glm(num_low_birthweight_births ~ eqi, 
          data=eqi_lbw_clean_df, 
          family=poisson, 
          offset=log_live_births)

summary(mod1_p)

sum(resid(mod1_p,type="pearson")^2)/mod1_p$df.residual
#yes dispersion is potential problem bc scale > 1

#goodness of fit
pchisq(mod1_p$deviance, mod1_p$df.residual, lower.tail=F)
#seeing a lack of fit for Poisson model

#Pearson residual plot
plot(resid(mod1_p,type="pearson"),ylab="Pearson residuals")
#also potential issue with outliers

```

#Prelim model just with EQI as exposure and low bw as outcome
##Quasipoisson model
```{r}
######Fit a Quasipoisson model (which acounts for overdispersion)
#fit the Poisson model (accounting for overdispersion)
#no offset bc exposure unit is already same here
mod1_qp = glm(num_low_birthweight_births ~ eqi, 
          data=eqi_lbw_clean_df, 
          family=quasipoisson, 
          offset=log_live_births)

summary(mod1_qp)

sum(resid(mod1_qp,type="pearson")^2)/mod1_qp$df.residual
#yes dispersion 

#goodness of fit
pchisq(mod1_qp$deviance, mod1_qp$df.residual, lower.tail=F)
#seeing a lack of fit with the quasipoisson

#Pearson residual plot
plot(resid(mod1_qp,type="pearson"),ylab="Pearson residuals")
#also potential issue with outliers

plot(mod1_qp)
#looks perhaps nonlinear?

```

#Prelim model just with EQI as exposure and low bw as outcome
##Penalized Spline
```{r}
mod1_qp_nl <- gam(num_low_birthweight_births ~ s(eqi),
                  family = "quasipoisson",
                  offset=log_live_births,
                  data = eqi_lbw_clean_df)

summary(mod1_qp_nl)

#this is the penalty estimated by the model 
mod1_qp_nl$sp

plot(mod1_qp_nl)

```
Seeing nonlinear relationship between eqi and num_low_birthweight_births

#Nonlinear checks of other indices and the low bw outcome; unadjusted
```{r}
##SOCIAL
mod2_qp_nl <- gam(num_low_birthweight_births ~ s(sociod), 
                  family = "quasipoisson",
                  offset=log_live_births,
                  data = eqi_lbw_clean_df)

summary(mod2_qp_nl)
plot(mod2_qp_nl)
#nonlinear

##AIR
mod3_qp_nl <- gam(num_low_birthweight_births ~ s(air), 
                  family = "quasipoisson",
                  offset=log_live_births,
                  data = eqi_lbw_clean_df)

summary(mod3_qp_nl)
plot(mod3_qp_nl)
#nonlinear

##BUILT
mod4_qp_nl <- gam(num_low_birthweight_births ~ s(built), 
                  family = "quasipoisson",
                  offset=log_live_births,
                  data = eqi_lbw_clean_df)

summary(mod4_qp_nl)
plot(mod4_qp_nl)
#nonlinear

##Water
mod5_qp_nl <- gam(num_low_birthweight_births ~ s(water), 
                  family = "quasipoisson",
                  offset=log_live_births,
                  data = eqi_lbw_clean_df)

summary(mod5_qp_nl)
plot(mod5_qp_nl)
#nonlinear
#water index was weird distribution; establish a cutoff?

##Land
mod6_qp_nl <- gam(num_low_birthweight_births ~ s(land), 
                  family = "quasipoisson",
                  offset=log_live_births,
                  data = eqi_lbw_clean_df)

summary(mod6_qp_nl)
plot(mod6_qp_nl)
#nonlinear

```

#Are there nonlinear relationships between the 5 subdomain indices?
##penalized spline
#using built index as the outcome for now
```{r}
##SOCIAL
mod7_qp_nl <- gam(built ~ s(sociod),
                  data = eqi_lbw_clean_df)

summary(mod7_qp_nl)
plot(mod7_qp_nl)
#nonlinear

##AIR
mod8_qp_nl <- gam(built ~ s(air), 
                  data = eqi_lbw_clean_df)

summary(mod8_qp_nl)
plot(mod8_qp_nl)
#nonlinear

##Water
mod9_qp_nl <- gam(built ~ s(water), 
                  data = eqi_lbw_clean_df)

summary(mod9_qp_nl)
plot(mod9_qp_nl)
#nonlinear; again water has weird distribution; use cutoff?

##Land
mod10_qp_nl <- gam(built ~ s(land), 
                  data = eqi_lbw_clean_df)

summary(mod10_qp_nl)
plot(mod10_qp_nl)
#very nonlinear

```

```

