---
title: "Method3_PCA"
output: html_document
date: 2025-04-02
---

# Set Up

## Load packages

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(readr)
library(readxl)
library(ggcorrplot)
library(janitor)
library(factoextra)
library(Qtools)
library(quantreg)
library(patchwork)


```

## Setting template

```{r}
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "right"))
 
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

## Load Datasets
```{r}
#main dataset
eqi_lbw_clean_df <- read_csv("data/eqi_lbw_clean_df.csv")
head(eqi_lbw_clean_df)
dim(eqi_lbw_clean_df)

#loading df with names of components that comprise each subdomain 
pca_input_names_df <- read_xlsx(path = "data/Data_Dictionary_Variables_EQI_2006_2010.xlsx", 
                                sheet = 1) |> 
  clean_names(case = "snake")

pca_input_names_df$variable_name <- tolower(pca_input_names_df$variable_name)

#loading df with transformed values of components that comprise each subdomain
pca_input_variables_transformed_raw <- read_csv("data/PCA_Input_Variables.csv")

```

## Clean PCA Input Variables Transformed Dataset
```{r}
#Add leading zeros to the stFIPS; code taken from the provided EQI README file 
pca_input_variables_transformed_raw$stfips <- substr(as.numeric(pca_input_variables_transformed_raw$stfips) + 100000, 2,7)

#rename variables
pca_input_variables_transformed_raw <- pca_input_variables_transformed_raw |> rename(fips = stfips)

#include only FIPS included in our main dataset
pca_input_variables_transformed_df <- pca_input_variables_transformed_raw |> filter(fips %in% eqi_lbw_clean_df$fips)

names(pca_input_variables_transformed_df) <- tolower(names(pca_input_variables_transformed_df))

head(pca_input_variables_transformed_df)
dim(pca_input_variables_transformed_df)
colnames(pca_input_variables_transformed_df)

```

## Matrix for Air
```{r}
#filter for only for air subdomain components
air_input_var_names <- pca_input_names_df |> filter(domain == "Air") 
dim(air_input_var_names)
#43 components in air domain

#create df with only air components
air_df <- pca_input_variables_transformed_df |> 
  select(fips, all_of(air_input_var_names$variable_name))

#create air matrix for pca
##rows are participants, columns are air components
##define rownames of matrix before running PCA
air_pca_input_matrix <- air_df |> 
  column_to_rownames("fips") |> 
  as.matrix()

```

## Air Spearman Correlations
```{r}
air_corr_inputs <- air_df |> select(-fips)

##Spearman correlations
air_corr <- round(cor(air_corr_inputs, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(air_corr_inputs)

#simple plot
ggcorrplot(air_corr)

#hiearchial clustering
ggcorrplot(air_corr, hc.order = TRUE, outline.color = "white")

#lower triangle
air_cor_plot <- ggcorrplot(air_corr,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white") + 
  ggtitle("Air Domain") 


#add correlation values onto plot
ggcorrplot(air_corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(air_corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)
```

## Air PCA
```{r}
#PCA, center and scale data
air_pca <- prcomp(air_pca_input_matrix, center = TRUE, scale = TRUE) 
summary(air_pca)

```

## Air Eigenvalues
```{r}
#Air eigenvalues
get_eig(air_pca)

#Air screeplot; percentage of total variance explained by each PC
air_screeplot <- fviz_eig(air_pca, main = "Air PCA",
         xlab = "Principal component", addlabels=TRUE) +
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Add 10% space on top



```
Air PC1 explains 28.1% of the total variation in the Air subdomain.

## Air PCA Scores
```{r}
#PCA scores pertain to participants; these are the coordinates of our samples on the new PC axis
#The PC scores are stored in the "x" value of the prcomp object
#Extract air pc scores
air_pc_scores <- air_pca$x

#make df
#rows are participants, columns have participants score within each PC
air_pc_scores_df <- air_pc_scores |> 
  as_tibble(rownames = "fips") #to go back to df format, convert rowname into column name

nrow(air_pc_scores_df)
head(air_pc_scores_df)

#plot of PC1 by PC2
air_pc_scores_df |>  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point()
#seeing one big cluster of points; PCs are not grouping separately

```

## Air Loadings; examine which components have the most influence
```{r}
#extract loadings; pertains to components
air_pc_loadings <- air_pca$rotation

#make df
air_pc_loadings_df <- air_pc_loadings |> 
  as_tibble(rownames = "air_components")

#Graph of All loadings in PC1
air_pc_loadings_df |> 
  ggplot(aes(x = reorder(air_components, PC1), y = PC1)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Air PC1 Loadings")

#Top 3 loadings with greatest magnitude in PC1
air_top3 <- air_pc_loadings_df |> 
  arrange(desc(abs(PC1))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(air_components, -abs(PC1)), y = PC1)) + # ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Air PC1: Top 3 Absolute Loadings",
       x = "Air Domain") +
  scale_x_discrete(labels = c("a_ebenzine_ln" = "Ethylbenzene",
                              "a_c2hcl3_ln" = "Trichloroethylene",
                              "a_co_ln" = "Cobalt compounds"))

```
The top 3 absolute loadings for Air PC1 are a_ebenzine_ln (Ethylbenzene), a_c2hcl3_ln (Trichloroethylene), and a_co_ln (Cobalt compounds). These components are contributing the most to the variation captured in Air PC1.

## Matrix for Built
```{r}
#filter for only for built subdomain components
built_input_var_names <- pca_input_names_df |> filter(domain == "Built") 
dim(built_input_var_names)
#15 components in built domain

#create df with only built components
built_df <- pca_input_variables_transformed_df |> 
  select(fips, all_of(built_input_var_names$variable_name))

#create built matrix for pca
##rows are participants, columns are air components
##define rownames of matrix before running PCA
built_pca_input_matrix <- built_df |> 
  column_to_rownames("fips") |> 
  as.matrix()

```

## Built Spearman Correlations
```{r}
built_corr_inputs <- built_df |> select(-fips)

##Spearman correlations
built_corr <- round(cor(built_corr_inputs, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(built_corr_inputs)

#simple plot
ggcorrplot(built_corr)

#hiearchial clustering
ggcorrplot(built_corr, hc.order = TRUE, outline.color = "white")

#lower triangle
built_cor_plot <- ggcorrplot(built_corr,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white") + 
  ggtitle("Built Domain") 

#add correlation values onto plot
ggcorrplot(built_corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(built_corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)
```

## Built PCA
```{r}
#PCA, center and scale data
built_pca <- prcomp(built_pca_input_matrix, center = TRUE, scale = TRUE) 
summary(built_pca)

```

## Built Eigenvalues
```{r}
#Built eigenvalues
get_eig(built_pca)

#Built screeplot; percentage of total variance explained by each PC
built_screeplot <- fviz_eig(built_pca, main = "Built PCA",
         xlab = "Principal component", addlabels=TRUE) + 
  theme(plot.title = element_text(hjust = 0.5)) +
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Add 10% space on top



```
Built PC1 explains 25.3% of the total variation in the Built subdomain.

## Built PCA Scores
```{r}
#PCA scores pertain to participants
#Extract built pc scores
built_pc_scores <- built_pca$x

#make df
#rows are participants, columns have participants score within each PC
built_pc_scores_df <- built_pc_scores |> 
  as_tibble(rownames = "fips") #to go back to df format, convert rowname into column name

nrow(built_pc_scores_df)
head(built_pc_scores_df)

#plot of PC1 by PC2
built_pc_scores_df |>  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point()
#seeing one big cluster of points; PCs are not grouping separately

```

## Built Loadings; examine which components have the most influence
```{r}
#extract loadings; pertains to components
built_pc_loadings <- built_pca$rotation

#make df
built_pc_loadings_df <- built_pc_loadings |> 
  as_tibble(rownames = "built_components")

#Graph of All loadings in PC1
built_pc_loadings_df |> 
  ggplot(aes(x = reorder(built_components, PC1), y = PC1)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Built PC1 Loadings")

#Top 3 loadings with greatest magnitude in PC1
built_top3 <- built_pc_loadings_df |> 
  arrange(desc(abs(PC1))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(built_components, -abs(PC1)), y = PC1)) + # ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Built PC1: Top 3 Absolute Loadings",
       x = "Built Domain") +
  scale_x_discrete(labels = c("ss_env_rate_ln" = "Rate Social Service\nAgences",
                              "rec_env_rate_ln" = "Rate Recreation-Related\nBusinesses",
                              "ed_env_rate_ln" = "Rate Education-Related\nBusinesses"))


```
The top 3 absolute loadings for Built PC1 are ss_env_rate_ln (Rate of social service agencies per county), rec_env_rate_ln (Rate of recreation-related businesses per county), and ed_env_rate_ln (Rate of education-related business per county). These components are contributing the most to the variation captured in Built PC1.

## Matrix for Land
```{r}
#filter for only for land subdomain components
land_input_var_names <- pca_input_names_df |> filter(domain == "Land") 
dim(land_input_var_names)
#18 components in land domain

#create df with only land components
land_df <- pca_input_variables_transformed_df |> 
  select(fips, all_of(land_input_var_names$variable_name))

#create land matrix for pca
##rows are participants, columns are air components
##define rownames of matrix before running PCA
land_pca_input_matrix <- land_df |> 
  column_to_rownames("fips") |> 
  as.matrix()

```

## Land Spearman Correlations
```{r}
land_corr_inputs <- land_df |> select(-fips)

##Spearman correlations
land_corr <- round(cor(land_corr_inputs, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(land_corr_inputs)

#simple plot
ggcorrplot(land_corr)

#hiearchial clustering
ggcorrplot(land_corr, hc.order = TRUE, outline.color = "white")

#lower triangle
land_cor_plot <- ggcorrplot(land_corr,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white") + 
  ggtitle("Land Domain") 


#add correlation values onto plot
ggcorrplot(land_corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(land_corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)
```

## Land PCA
```{r}
#PCA, center and scale data
land_pca <- prcomp(land_pca_input_matrix, center = TRUE, scale = TRUE) 
summary(land_pca)

```

## Land Eigenvalues
```{r}
#Land eigenvalues
get_eig(land_pca)

#Land screeplot; percentage of total variance explained by each PC
land_screeplot <- fviz_eig(land_pca, main = "Land PCA",
         xlab = "Principal component", addlabels=TRUE) + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Add 10% space on top

```
Land PC1 explains 27.7% of the total variation in the Land subdomain. PC2 explains 14.2% of total variation in Land subdomain.

## Land PCA Scores
```{r}
#PCA scores pertain to participants
#Extract built pc scores
land_pc_scores <- land_pca$x

#make df
#rows are participants, columns have participants score within each PC
land_pc_scores_df <- land_pc_scores |> 
  as_tibble(rownames = "fips") #to go back to df format, convert rowname into column name

nrow(land_pc_scores_df)
head(land_pc_scores_df)

#plot of PC1 by PC2
land_pc_scores_df |>  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point()
#seeing some clustering for PC2

```

## Land Loadings; examine which components have the most influence
```{r}
#extract loadings; pertains to components
land_pc_loadings <- land_pca$rotation

#make df
land_pc_loadings_df <- land_pc_loadings |> 
  as_tibble(rownames = "land_components")

#Graph of All loadings in PC1
land_pc_loadings_df |> 
  ggplot(aes(x = reorder(land_components, PC1), y = PC1)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Land PC1 Loadings")

#Top 3 loadings with greatest magnitude in PC1
land_top3 <- land_pc_loadings_df |> 
  arrange(desc(abs(PC1))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(land_components, -abs(PC1)), y = PC1)) + #ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Land PC1: Top 3 Absolute Loadings",
       x = "Land Domain") +
    scale_x_discrete(labels = c("pct_harvested_acres_ln" = "% Harvested\nAcres",
                              "pct_disease_acres_ln" = "% Crops Treated\nwith Chemicals",
                              "pct_manure_acres_ln" = "% Manure\n Per Acre"))


#Graph of All loadings in PC2
land_pc_loadings_df |> 
  ggplot(aes(x = reorder(land_components, PC2), y = PC2)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Land PC2 Loadings")

#Top 3 loadings with greatest magnitude in PC2
land_pc_loadings_df |> 
  arrange(desc(abs(PC2))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(land_components, -abs(PC2)), y = PC2)) + #ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Land PC2: Top 3 Absolute Loadings")

```
The top 3 absolute loadings for Land PC1 are pct_harvested_acres_ln (Harvested acres, acres harvested per county acres), pct_disease_acres_ln (Chemicals used to control Diseases in crops and orchards, acres applied per county acres), and pct_manure_acres_ln (Manure, acres applied per county acres). These components are contributing the most to the variation captured in Land PC1.
The top 3 absolute loadings for Land PC2 are insecticide_ln (Insecticide applied), fungicide_ln (Fungicides applied), and pct_au_ln (Animal Units, animal units per county acres). These components are contributing the most to the variation captured in Land PC2.

## Matrix for Sociodemographic
```{r}
#filter for only for built subdomain components
sociod_input_var_names <- pca_input_names_df |> filter(domain == "Sociodemographic") 
dim(sociod_input_var_names)
#12 components in sociod domain

#create df with only sociod components
sociod_df <- pca_input_variables_transformed_df |> 
  select(fips, all_of(sociod_input_var_names$variable_name))

#create sociod matrix for pca
##rows are participants, columns are air components
##define rownames of matrix before running PCA
sociod_pca_input_matrix <- sociod_df |> 
  column_to_rownames("fips") |> 
  as.matrix()

```

## Sociod Spearman Correlations
```{r}
sociod_corr_inputs <- sociod_df |> select(-fips)

##Spearman correlations
sociod_corr <- round(cor(sociod_corr_inputs, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(sociod_corr_inputs)

#simple plot
ggcorrplot(sociod_corr)

#hiearchial clustering
ggcorrplot(sociod_corr, hc.order = TRUE, outline.color = "white")

#lower triangle
sociod_cor_plot <- ggcorrplot(sociod_corr,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white") + 
  ggtitle("Sociodemographic Domain") 

#add correlation values onto plot
ggcorrplot(sociod_corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(sociod_corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)
```

## Sociod PCA
```{r}
#PCA, center and scale data
sociod_pca <- prcomp(sociod_pca_input_matrix, center = TRUE, scale = TRUE) 
summary(sociod_pca)

```

## Sociod Eigenvalues
```{r}
#Sociod eigenvalues
get_eig(sociod_pca)

#Built screeplot; percentage of total variance explained by each PC
sociod_screeplot <- fviz_eig(sociod_pca, main = "Sociod PCA",
         xlab = "Principal component", addlabels=TRUE) + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Add 10% space on top

```
Sociod PC1 explains 29.9% of the total variation in the Sociod subdomain.
Sociod PC2 explains 22% of the total variation in the Sociod subdomain.

## Sociod PCA Scores
```{r}
#PCA scores pertain to participants
#Extract built pc scores
sociod_pc_scores <- sociod_pca$x

#make df
#rows are participants, columns have participants score within each PC
sociod_pc_scores_df <- sociod_pc_scores |> 
  as_tibble(rownames = "fips") #to go back to df format, convert rowname into column name

nrow(sociod_pc_scores_df)
head(sociod_pc_scores_df)

#plot of PC1 by PC2
sociod_pc_scores_df |>  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point()
#seeing some separation but no distinct clusters

```

## Sociod Loadings; examine which components have the most influence
```{r}
#extract loadings; pertains to components
sociod_pc_loadings <- sociod_pca$rotation

#make df
sociod_pc_loadings_df <- sociod_pc_loadings |> 
  as_tibble(rownames = "sociod_components")

#Graph of All loadings in PC1
sociod_pc_loadings_df |> 
  ggplot(aes(x = reorder(sociod_components, PC1), y = PC1)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Sociod PC1 Loadings")

#Top 3 loadings with greatest magnitude in PC1
sociod_top3 <- sociod_pc_loadings_df |> 
  arrange(desc(abs(PC1))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(sociod_components, -abs(PC1)), y = PC1)) + # ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Sociod PC1: Top 3 Absolute Loadings",
       x = "Sociodemographic Domain") +
  scale_x_discrete(labels = c("num_creatclass" = "% Employed in\nCreative Class",
                              "pct_bs" = "% Bachelor's Degree\nor Higher",
                              "med_hh_value" = "Median Household\nValue"))


#Graph of All loadings in PC2
sociod_pc_loadings_df |> 
  ggplot(aes(x = reorder(sociod_components, PC2), y = PC2)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Sociod PC2 Loadings")

#Top 3 loadings with greatest magnitude in PC2
sociod_pc_loadings_df |> 
  arrange(desc(abs(PC2))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(sociod_components, -abs(PC2)), y = PC2)) + # ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Sociod PC2: Top 3 Absolute Loadings")
```
The top 3 absolute loadings for Sociod PC1 are num_creatclass (Percent county employed in creative class), pct_bs (Bachelor's degree or higher, percent of persons age 25 years+), and med_hh_value (Median household value). These components are contributing the most to the variation captured in Sociod PC1.

The top 3 absolute loadings for Sociod PC2 are gini_est (Measure of income inequality), pct_renterocc (Percent renter-occupied units), and pct_fam_pov (Percent of families in poverty). These components are contributing the most to the variation captured in Sociod PC2.

## Matrix for Water
```{r}
#filter for only for built subdomain components
water_input_var_names <- pca_input_names_df |> filter(domain == "Water") 
dim(water_input_var_names)
#51 components in sociod domain

#create df with only water components
water_df <- pca_input_variables_transformed_df |> 
  select(fips, all_of(water_input_var_names$variable_name))

#create water matrix for pca
##rows are participants, columns are air components
##define rownames of matrix before running PCA
water_pca_input_matrix <- water_df |> 
  column_to_rownames("fips") |> 
  as.matrix()
dim(water_pca_input_matrix)

```

## Water Spearman Correlations
```{r}
water_corr_inputs <- water_df |> select(-fips)

#Since Spearman correlation is based on ranks, a variable with zero variance (no variation) can't be ranked meaningfully.

#Identify columns with zero variance (constant columns)
water_zero_variance_cols <- apply(water_corr_inputs, 2, function(x) sd(x) == 0)
#2 indicates that the function is applied to columns

#double check
sd(water_corr_inputs$w_ethylbenz_ln)

#Remove columns with zero variance from the df; here it is w_ethylbenz_ln
water_corr_inputs_clean <- water_corr_inputs[, !water_zero_variance_cols]
dim(water_corr_inputs_clean)

##Spearman correlations
water_corr <- round(cor(water_corr_inputs_clean, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(water_corr_inputs_clean)

#simple plot
ggcorrplot(water_corr)

#hierarchical clustering
ggcorrplot(water_corr, hc.order = TRUE, outline.color = "white")

#lower triangle
water_cor_plot <- ggcorrplot(water_corr,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white") + 
  ggtitle("Water Domain") 


#add correlation values onto plot
ggcorrplot(water_corr,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(water_corr,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)
```

## Water PCA
```{r}
#PCA, center and scale data
#water_pca <- prcomp(water_pca_input_matrix, center = TRUE, scale = TRUE) 
#summary(water_pca)

#receiving error message: Error in prcomp.default(water_pca_input_matrix, center = TRUE, scale = TRUE) cannot rescale a constant/zero column to unit variance

#Identify columns with zero variance (constant columns)
zero_variance_cols <- apply(water_pca_input_matrix, 2, function(x) sd(x) == 0)
#2 indicates that the function is applied to columns

#double check
sd(water_df$w_ethylbenz_ln)
sd(water_df$w_chlordane_ln)

##Why are w_ethylbenz_ln values are the same??

#Remove columns with zero variance from the df
water_pca_input_matrix_clean <- water_pca_input_matrix[, !zero_variance_cols]
#saying if zero_variance_cols is TRUE then don't include in the dataset
dim(water_pca_input_matrix_clean)
#water_pca_input_matrix_clean[, "w_ethylbenz_ln"]

#PCA, center and scale data
water_pca <- prcomp(water_pca_input_matrix_clean, center = TRUE, scale = TRUE) 
summary(water_pca)

#now using only 50 components for water
```

## Water Eigenvalues
```{r}
#Water eigenvalues
get_eig(water_pca)

#Built screeplot; percentage of total variance explained by each PC
water_screeplot <- fviz_eig(water_pca, main = "Water PCA",
         xlab = "Principal component", addlabels=TRUE) + 
  theme(plot.title = element_text(hjust = 0.5)) + 
  scale_y_continuous(expand = expansion(mult = c(0, 0.1)))  # Add 10% space on top


```
Water PC1 explains 42% of the total variation in the Water subdomain.
Water PC2 explains 8.9% of the total variation in the Water subdomain.

## Water PCA Scores
```{r}
#PCA scores pertain to participants
#Extract built pc scores
water_pc_scores <- water_pca$x

#make df
#rows are participants, columns have participants score within each PC
water_pc_scores_df <- water_pc_scores |> 
  as_tibble(rownames = "fips") #to go back to df format, convert rowname into column name

nrow(water_pc_scores_df)
head(water_pc_scores_df)

#plot of PC1 by PC2
water_pc_scores_df |>  
  ggplot(aes(x = PC1, y = PC2)) +
  geom_point()
#seeing some clustering

```

## Water Loadings; examine which components have the most influence
```{r}
#extract loadings; pertains to components
water_pc_loadings <- water_pca$rotation

#make df
water_pc_loadings_df <- water_pc_loadings |> 
  as_tibble(rownames = "water_components")

#Graph of All loadings in PC1
water_pc_loadings_df |> 
  ggplot(aes(x = reorder(water_components, PC1), y = PC1)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Water PC1 Loadings")

#Top 3 loadings with greatest magnitude in PC1
water_top3 <- water_pc_loadings_df |> 
  arrange(desc(abs(PC1))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(water_components, -abs(PC1)), y = PC1)) + # ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Water PC1: Top 3 Absolute Loadings",
       x = "Water Domain") +
  scale_x_discrete(labels = c("w_c2cl4_ln" = "Tetrachloroethylene",
                              "w_trichlorene_ln" = "Trichloroethylene",
                              "w_benzene_ln" = "Monochlorobenzene"))


#Graph of All loadings in PC2
water_pc_loadings_df |> 
  ggplot(aes(x = reorder(water_components, PC2), y = PC2)) + 
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_hline(yintercept = 0) +
  labs(title = "Water PC2 Loadings")

#Top 3 loadings with greatest magnitude in PC2
water_pc_loadings_df |> 
  arrange(desc(abs(PC2))) |> 
  slice(1:3) |> 
  ggplot(aes(x = reorder(water_components, -abs(PC2)), y = PC2)) + # ordered by descending absolute values
  geom_col() +
  theme(legend.position = "bottom", axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title = "Water PC2: Top 3 Absolute Loadings")
```
The top 3 absolute loadings for Water PC1 are w_c2cl4_ln (Tetrachloroethylene), w_trichlorene_ln (Trichloroethylene), and w_benzene_ln (Monochlorobenzene (Chlorobenzene)) These components are contributing the most to the variation captured in Water PC1.

The top 3 absolute loadings for Water PC2 are w_picloram_ln (Picloram), w_dinoseb_ln (Dinoseb), and w_endrin_ln (Endrin). These components are contributing the most to the variation captured in Water PC2.

# Quick Look at Correlations (using Spearman) for all pca input variables
```{r}
pca_inputs <- pca_input_variables_transformed_df[,5:143]
dim(pca_inputs)

#Since Spearman correlation is based on ranks, a variable with zero variance (no variation) can't be ranked meaningfully.

#Identify columns with zero variance (constant columns)
pca_inputs_zero_variance_cols <- apply(pca_inputs, 2, function(x) sd(x) == 0)
#2 indicates that the function is applied to columns

#double check
sd(pca_inputs$w_ethylbenz_ln)

#Remove columns with zero variance from the df; here it is w_ethylbenz_ln
pca_inputs_clean <- pca_inputs[, !pca_inputs_zero_variance_cols]
dim(pca_inputs_clean)

##Spearman correlations
corr_pca <- round(cor(pca_inputs_clean, method = "spearman"), 2)

#calculate p-values for spearman correlations
p_mat <- cor_pmat(pca_inputs_clean)

#simple plot
ggcorrplot(corr_pca)

#hiearchial clustering
ggcorrplot(corr_pca, hc.order = TRUE, outline.color = "white")

#lower triangle
ggcorrplot(corr_pca,
           hc.order = TRUE,
           type = "lower",
           outline.color = "white")

#add correlation values onto plot
ggcorrplot(corr_pca,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE)

#add correlation significance level
#X marks the not signif correlations (Barring the no significant coefficient)
ggcorrplot(corr_pca,
           hc.order = TRUE,
           type = "lower",
           p.mat = p_mat)

```

#Regression
```{r}
air_PC1_df <- air_pc_scores_df |> select(fips,PC1) |> rename(pc1_air = PC1)
built_PC1_df <- built_pc_scores_df |> select(fips,PC1) |> rename(pc1_built = PC1)
land_PC1_df <- land_pc_scores_df |> select(fips,PC1)|> rename(pc1_land = PC1)
sociod_PC1_df <- sociod_pc_scores_df |> select(fips,PC1)|> rename(pc1_sociod = PC1)
water_PC1_df <- water_pc_scores_df |> select(fips,PC1) |> rename(pc1_water = PC1)

eqi_lbw_clean_df <- eqi_lbw_clean_df |> 
  mutate(cat_rucc = factor(cat_rucc, levels = c("metropolitan-urbanized", "less urbanized", "thinly populated", "nonmetropolitan-urbanized")))

a <- left_join(eqi_lbw_clean_df, air_PC1_df, by = "fips")
b <- left_join(a, built_PC1_df, by = "fips")
c <- left_join(b, land_PC1_df, by = "fips")
d <- left_join(c, sociod_PC1_df, by = "fips")
eqi_lbw_clean_pca_df <- left_join(d, water_PC1_df, by = "fips")

mod = glm(num_low_birthweight_births ~ pc1_air + pc1_built + pc1_land + pc1_sociod + pc1_water + 
             cat_rucc + perc_age1849 + perc_smokers, 
          data = eqi_lbw_clean_pca_df, 
          family = "quasipoisson", 
          offset = log_live_births)

summary(mod)

```
We cannot feasibly rerun the regression using our self-generated PC1s, because the EPA used a valence correction that we are unable to replicate. Thus, without applying the valence correction, the direction of our relationships are not consistent with EPA provided PC1s.

#Figure with all correlations
```{r}
air_cor_plot + built_cor_plot + land_cor_plot + sociod_cor_plot + water_cor_plot
```

#Figure with all elbow plots
```{r}
air_screeplot + built_screeplot + land_screeplot + sociod_screeplot + water_screeplot
```


#Figure with all 3 top loadings
```{r}
air_top3 + built_top3 + land_top3 + sociod_top3 + water_top3

```

